---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
You need to write code to answer each question. No credit for just searching through the dataset.

1. Consider the dataset faithful. What is the mean waiting for the observations where the eruption is longer than 2 minutes.
```{r}

mean(faithful[faithful$eruptions > 2, 'waiting'])

```

2. Consider the dataset state.x77. 
This dataset is actually a _______ so we will create a data_frame from it.
```{r}
## series of vectors that contain data
state2<-as.data.frame(state.x77)
str(state2)
state2

```

a. What variables does state2 have in it and what are their types?
b. What is the median income for the states?
c. What is the mean income for the states?
d. What is the range of  the lowest 25% of the incomes?
e. Graph the histogram of incomes. Describe the shape of the histogram.
```{r}
## a.
#var_state2 <- names(state2)
#var_state2
#typeof(var_state2)
str(state2)

## b.
median(state2$Income)

## c.
mean(state2$Income)

## d.

q1=fivenum(state2$Income,na.rm=T)[2]
range(state2$Income[state2$Income < q1])


## e.

incomes <- state2$Income
hist(incomes)

##From the histogram it seems that the shape seems to be bell shaped since 
##it has one peak in the middle of the graph.





```


3. There are  other  variables for state: state.region, state.name, state.abb.
a. What regions are there?
b. What states are in the region west?
c. What is the median HS graduation rate for the states that are in the south?
d. What state in the west has the smallest life expectancy?
e. Create a table with the number of states in each regions  with populations greater than 5000 K and the number in each region  whose population is not.
```{r}
## a.
levels(state.region)

## b.
state.name[state.region == 'West']

## c.
median(state2$`HS Grad`[state.region == 'South'])

## d.
min(state2$`Life Exp`[state.region == 'West'])

## e.

table(state.region,state2$Population > 5000)


```

4.Plot discoveries.What do you observe? What might be an interesting plot to graph simultaneously?
```{r}
plot(discoveries)
## I observe that the plot is telling me the number of inventions from 
#the years 1860 to 1959. The frequency tells me the number of inventions per year.
#For example, for 1860 there were 5 inventions, for 1861 there were 3 inventions, 
#and so on. The plot seems to have spike in discoveries in the early 1880's.
#An interesting plot to graph simultaneously would be uspop. 
#We can see if there is any connection between the US population and the 
#number of inventions in a time period.
```



5. a.Graph the boxplot of unicef totpop.You will need to include code to load the unicef dataset. Are there outliers? Ignoring the outliers, what do you observe? 
b. What is the five number summary for totpop?
```{r}
## a.
load("C:/Users/akshe/Desktop/unicefbas2016 (1).rda")
boxplot(unicefb2016$totpop, xlab = c("totpop"))

## I have noticed that there are outliers. 
#One at 1.3 million and one at 1.4 million. Ignoring the outliers,
#I have noticed that the total populations are quite low at about 7500. 

## b.
summary(unicefb2016$totpop)



```


Knit your file and submit it to Canvas as a pdf.
